Cross-Loop Optimization of Arithmetic Intensity for Finite Element Local Assembly

### 2014-10-20 TACO-paper-v2/coffee.tex
Reapply Ram's edits
edits

We aim to maximize our impact on the platforms that are realistically used
for finite element applications, so we target conventional CPU architectures
rather than GPUs. 
The key limiting factor to the execution on GPUs is the stringent memory requirements. 
Only relatively small problems fit in a GPU memory, and support for
distributed GPU execution in general purpose finite element frameworks is minimal. There has been some research on
adapting local assembly to GPUs, which we mention in Section~\ref{sec:related-work},
although it differs from ours in several ways, including: (i) not relying on
automated code generation from a domain-specific language (explained next),
(ii) testing only order 1 methods (while we focus on the range 1-4), (iii) not optimizing for
cross-loop arithmetic intensity (the goal is rather effective multi-thread
parallelization). In addition, our code transformations would drastically impact the 
GPU parallelization strategy, for example by increasing a thread's working set. 
For all these reasons, a study on extending our work to GPU architectures is 
beyond the scope of this paper. In Section~\ref{sec:generality}, however, we provide 
some intuitions about this research direction.

@@

Splitting an expression (henceforth \emph{split}) has, however, several drawbacks. Firstly, it
increases the number of accesses to $A$ proportionally to the ``split
factor'', which is the number of sub-expressions produced. Also,
depending on how the split is executed, it can lead to redundant
computation. For example, the number of times the product $det*W3[i]$ is
performed is proportional to the number of sub-expressions, as shown in
the code snippet. Further, it increases loop overhead, for example
through additional branch instructions. Finally, it might affect
register locality: for instance, the same array could be accessed in
different sub-expressions, requiring a proportional number of loads be
performed. This is not the case for the Helmholtz example. Nevertheless,
as shown in Section~\ref{sec:perf-results}, the performance gain from
improved register reuse along inner dimensions can still be greater,
especially if the split factor and the splitting itself use heuristics
to minimize the aforementioned issues.
~\\

>>>>

We aim to maximize our impact on the platforms that are realistically used
for finite element applications, so we target conventional CPU architectures
rather than GPUs. 
The key limiting factor to the execution on GPUs is the stringent
memory requirements.  Only relatively small problems fit in a GPU
memory, and support for distributed GPU execution in general purpose
finite element frameworks is minimal. There has been some research on
adapting local assembly to GPUs, which we mention in
Section~\ref{sec:related-work}, although it differs from ours in
several ways, including: (i) not optimizing for cross-loop arithmetic
intensity (the goal is rather effective multi-thread parallelization),
and (ii) not relying on automated code generation from a
domain-specific language (explained next); moreover, they have done
testing only on order 1 methods, while we focus on the order range
1--4. In addition, our code transformations would drastically impact
the GPU parallelization strategy, for example by increasing a thread's
working set.  For all these reasons, a study on extending our work to
GPU architectures is beyond the scope of this paper. In
Section~\ref{sec:generality}, however, we provide some intuitions
about this research direction.

@@

Splitting an expression (henceforth \emph{split}) has, however, several drawbacks. Firstly, it
increases the number of accesses to $A$ proportionally to the ``split
factor,'' which is the number of sub-expressions produced. Also,
depending on how the split is executed, it can lead to redundant
computation. For example, the number of times the product $det*W3[i]$ is
performed is proportional to the number of sub-expressions, as shown in
the code snippet. Further, it increases loop overhead, for example
through additional branch instructions. Finally, it might affect
register locality: for instance, the same array could be accessed in
different sub-expressions, requiring a proportional number of loads be
performed. This is not the case for the Helmholtz example. Nevertheless,
as shown in Section~\ref{sec:perf-results}, the performance gain from
improved register reuse along inner dimensions can still be greater,
especially if the split factor and the splitting itself use heuristics
to minimize the aforementioned issues.
~\\
